---
title: "Chicago Crime Report"
author: "Blake Brown and Jacob Fullerton"
date: "9/26/2018"
output: html_document
knit: (function(input_file, encoding) {
  out_dir <- 'docs';
  rmarkdown::render(input_file,
 encoding=encoding,
 output_file=file.path(dirname(input_file), out_dir, 'index.html'))})
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Imported Packages
```{r Imports, message=FALSE, warning=FALSE}
library(tidyverse)
library(lubridate)
library(ggplot2)
library(ggmap)
library(dplyr)
library(data.table)
library(ggrepel)
library(varhandle)
library(gridExtra)
```

#Importing Data

To import all the data we read in from csv files each year range seperatly and the combine all the data with rbind to create a complete data set.

```{r Read CSV for data, warning=FALSE}
to2004 <- read.csv("Chicago_Crimes_2001_to_2004.csv",stringsAsFactors=FALSE)
to2007 <- read.csv("Chicago_Crimes_2005_to_2007.csv",stringsAsFactors=FALSE)
to2011 <- read.csv("Chicago_Crimes_2008_to_2011.csv",stringsAsFactors=FALSE)
to2017 <- read.csv("Chicago_Crimes_2012_to_2017.csv",stringsAsFactors=FALSE)
all <- rbind(to2004,to2007,to2011,to2017)
Chicago.Aft.2008 <- rbind(to2011, to2017)
```

In addition, to plot the data we must get a map of the area of Chicago
```{r Get Map, message=FALSE, warning=FALSE}
register_google(key = "AIzaSyC9rrtr993vzkQlEF3HfYdzwcj0ojOnLzA")
map <- get_map(location=c(lon=-87.645167,lat=41.808013), zoom=11, maptype='roadmap', color='bw')#Get the map from Google Maps
```


## Cleaning an sorting data

After the data has been imported we decided to only focus on the most recent 10 years to decrease processor load on the computer. Since some of the values were imported as strings they must be transformed to numeric values before they can be graphed. Finally, for the purpose of graphing only points with a latitude and longitude can be included so we created a new dataframe with only points with a Latitude and Longitude. We also decided to create new columns using the date to help with exploratory data analysis.

```{r Clean Data, message=FALSE, warning=FALSE}
crimes <- filter(all,Year>2007)
crimes <-filter(crimes,Year<2018)
crimes$Longitude <- as.numeric(crimes$Longitude)
crimes$Latitude <- as.numeric(crimes$Latitude)
hasLocation <- filter(crimes, !is.na(Longitude),!is.na(Latitude))
hasLocation$Day <- factor(day(as.POSIXlt(hasLocation$Date, format="%m/%d/%Y %I:%M:%S %p")))
hasLocation$Month <- factor(month(as.POSIXlt(hasLocation$Date, format="%m/%d/%Y %I:%M:%S %p")))
hasLocation$Year <- factor(year(as.POSIXlt(hasLocation$Date, format="%m/%d/%Y %I:%M:%S %p")))
hasLocation$Date <- as.Date(hasLocation$Date, "%m/%d/%Y %I:%M:%S %p")
Chicago.Aft.2008.Small <- Chicago.Aft.2008 %>% select(Date, Primary.Type, Location.Description, Arrest, Year, Latitude, Longitude)
Chicago.Aft.2008.Small <- na.omit(Chicago.Aft.2008.Small) ## removes any NA's

Chicago.Aft.2008.Small$Latitude <- as.numeric(Chicago.Aft.2008.Small$Latitude) ## Changes latitude to a numeric 
Chicago.Aft.2008.Small$Longitude <- as.numeric(Chicago.Aft.2008.Small$Longitude) ## Changes latitude to a numeric 
Chicago.Aft.2008.Small$Day <- factor(day(as.POSIXlt(Chicago.Aft.2008.Small$Date, format="%m/%d/%Y %I:%M:%S %p")))
Chicago.Aft.2008.Small$Month <- factor(month(as.POSIXlt(Chicago.Aft.2008.Small$Date, format="%m/%d/%Y %I:%M:%S %p")))
Chicago.Aft.2008.Small$Year <- factor(year(as.POSIXlt(Chicago.Aft.2008.Small$Date, format="%m/%d/%Y %I:%M:%S %p")))

Chicago.Aft.2008.Small$Date <- as.Date(Chicago.Aft.2008.Small$Date, "%m/%d/%Y %I:%M:%S %p")
Chicago.Aft.2008.Small <- Chicago.Aft.2008.Small[Chicago.Aft.2008.Small$Year != "2017", ] ##exclude the year 2017
Chicago.Aft.2008.Small <- na.omit(Chicago.Aft.2008.Small) ## removes any NA's introduced by coercion
```

## Exploratory Data Analysis
Before we go about looking at the distributions of the crimes we chose, it is important to get a better understanding of the data. In order to achieve this we will be making visualizations to help us understand the data we have. First we will look at the number of crimes.
```{r Exploratory Data Analysis, message=FALSE, warning=FALSE}
by_Crime <- hasLocation %>% group_by(Primary.Type) 
numCrimes <- summarize(by_Crime, num = n())
View(numCrimes)
numCrimes <- numCrimes[numCrimes$num>5000,]
## Histogram of the Crimes Committed
ggplot(numCrimes, aes(x=reorder(Primary.Type, num), num)) + geom_bar(stat = "identity", 
                                                                     color = "black", fill = "white") + 
  coord_flip() +
  xlab("Crime") + ylab("Number of Crimes Committed") + ggtitle("Distribution of Crimes Committed") +
  theme(
    axis.text.y = element_text(vjust = .5, hjust = 1, size = 5)
  )
```
From this graphic we can see that theft and battery are the most common crimes. One interesting note from this graphic would be the fact that there is a category called other offense. It would be interesting to see what is in this category. However, as it contains a wide variety of crimes it would be better to not use this in our analysis. Next we will explore the number of crimes committed based upon on the month.
```{r}
ggplot(hasLocation, aes(x=Month)) +
  geom_bar(colour="black", fill="white") +
  ylab('Number of Crimes Committed') + ggtitle("Distribution of Crimes per Month") +
  xlab("Month") + theme(plot.title = element_text(hjust = .5))
```
Interestingly, the number of crimes committed appears to peak during the summer months and sharply declines at the end of the year. This could be because during the summer months more people are out and about in Chicago and during the winter months most people elect to stay in their homes to avoid the cold. The more people being our causes more street and sidewalk crimes. One more interesting graphic to look at before we focus on only the crimes is the proportion of arresets by year and month.

```{r}
## Colors for the gradient scale
colors <- c('navyblue', 'darkmagenta', 'darkorange1')

## Gets only the rows that resulted in arrests
arrests_only <- Chicago.Aft.2008.Small[Chicago.Aft.2008.Small$Arrest == 'True',]
## Groups the arrests count by year and month with a total which is the count
by_arrests <- arrests_only %>% group_by(Year, Month) %>% summarise(Total = n())
## Groups by the Year and month with a total which is count
by_crime <- Chicago.Aft.2008.Small %>% group_by(Year, Month) %>% summarise(Total = n())
total <- cbind(crime_total=by_crime$Total, by_arrests)
total$prop <- total$Total/total$crime_total

prop <- ggplot(total, aes(Year, Month, fill = total$prop)) +
  geom_tile(size = 1, color = "white") + scale_fill_gradientn(colors = colors) + 
  geom_text(aes(label=round(total$prop, digits = 3)), color='white', size = 3) +
  labs(fill = "Proportion") +
  ggtitle("Proportion of Crimes Resulting in \nArrests by Year and Month") + 
  theme(legend.title = element_text(hjust = .5),  ## Centered the legend title 
        plot.title = element_text(hjust = .5), ## Centered the plot title
        panel.background = element_blank()
  )
prop

```
This graphic is very intersting. We can see that the mosts arrests occur around February every year. About 1/3 of the crimes committed in February resulted in an arrest. We can see that on average that about 26% percent of the crimes actually resulted in an arrest. This is much lower than what we would suspect it to be. This would suggest that there was not enough credible evidence or witnesses to support these crimes. 
```{r}
## Histogram of the number crimes committed for each location
by_Location <- Chicago.Aft.2008.Small %>% group_by(Location.Description) 
numLocations <- summarize(by_Location, num = n())
numLocations <- numLocations[numLocations$num>5000,]
ggplot(numLocations, aes(reorder(Location.Description,num), y = num)) + geom_bar(stat = "identity", color = "black", fill = "white") + coord_flip() +
  xlab("Location") + ylab("Number of Crimes Committed") + ggtitle("Distribution of Locations") +
  theme(
    axis.text.y = element_text(vjust = .5, hjust = 1, size = 5.5)
  )
```
We can see that the majority of crimes occur on the street and resential areas. Again there is an interesting "other" category. Again we will not include this in our analysis because there is no clear answer as to what "other" is. We will continue exploring by using only the six crimes we have chose, Narcotics, Assualt, Gambling, Kidnappung, Robbery, and Homicide.
```{r,message=FALSE, warning=FALSE}
## The primary crimes we chose
Primary_Crimes <- c("NARCOTICS", "ASSAULT", "GAMBLING", "KIDNAPPING", "ROBBERY","HOMICIDE")
## The data is only using our primary crimes
hasLocation <- hasLocation[hasLocation$Primary.Type %in% Primary_Crimes,]
## First we need to groupby the date
by_Date <- hasLocation%>% group_by(Date) %>% summarise(Total = n())
## Group by date and crime committed
by_Date2 <- hasLocation%>% group_by(Date, Primary.Type) %>% summarise(Total = n())
## Now plotting the graphic
ggplot(by_Date, aes(x=by_Date$Date, y=by_Date$Total)) + geom_smooth(se=FALSE) + 
  xlab("Date") + ylab("Total Amount of Crimes") + ggtitle("Number of Crimes per Day") +
  theme(plot.title = element_text(hjust = .5))
```
From this smooth line plot of the number of crimes per day we can see that crime has generally decreased from 2008, with a steep decline from 2010 to 2012. After 2012 we see that the line is generally decreasing, resulting in less crimes committed. 
```{r,message=FALSE, warning=FALSE}
## Coloring the lines based upon the 6 different crimes
ggplot(by_Date2, aes(x=by_Date2$Date, y=by_Date2$Total, color = by_Date2$Primary.Type)) + geom_smooth(se = FALSE) + 
  xlab("Date") + ylab("Total Amount of Crimes") + ggtitle("Number of Crimes per Day") +
  labs(color = "Crime") + theme(plot.title = element_text(hjust = .5),
                                legend.title = element_text(hjust = .5),
                                legend.position=c(1,1), legend.justification=c(1,1), 
                                legend.background = element_blank(),
                                legend.key = element_blank()
                                )
```

We can see that infact from 2008 to 2014 that most of the crimes decrease, expect those that were small amounts anyways. However, starting in 2014 we can see that the number assault, and robbery crimes have increased, passing the number of narcotic crimes. One theory for this is the focus on drug busting has cause other types of crime to become overlooked thus becoming committed more. Time to explore the number of crimes committed for each of the six crimes by month.

```{r,message=FALSE, warning=FALSE}
ggplot(hasLocation, aes(x=Month)) +
  geom_bar(colour="black", fill="white") +
  ylab('Count') + ggtitle("Distribution of Crimes per Month") +
  xlab("Month") + facet_wrap(~hasLocation$Primary.Type, scales='free')
```
This graphic is very intersting. One of the things that immediately pops out to me is the distribution for gambling. We can see that a majority of the gambling crimes are committed during the summer months and virtuallly non are committed during the winter months. We can also see that for assault, robbery, and homicide crimes the peak is during the summer. However, for kidnapping and narcotics the mose common months are late winter/early spring. It could be interesting to explore the location description to see if most of the crimes are committed outdoors, which would be why the distribution is like that.
```{r,message=FALSE, warning=FALSE,fig.width=5,fig.height=4}
## This is to help adjust the x axis values so the more popular crimes show only the more popular locations and
## the less popular crimes can show all the locations
by_Crime <- Chicago.Aft.2008.Small%>% group_by(Primary.Type, Location.Description) %>% summarize(Total = n())
by_Crime <- by_Crime[((by_Crime$Total > 2000 & by_Crime$Primary.Type %in% c("ASSAULT", "NARCOTICS", "ROBBERY")) | 
            (by_Crime$Total > 20 & by_Crime$Primary.Type %in% c("GAMBLING","HOMICIDE","KIDNAPPING") )),]

ggplot(by_Crime, aes(reorder(Location.Description, -Total), y=Total)) +
  geom_bar(stat = "identity",colour="black", fill="white") +
  ylab('Count') + ggtitle("Distribution of Crimes per Month") +
  xlab("Month") + 
  theme(axis.text.x = element_text(angle = 60, vjust = 1, hjust = 1, size = 6),
        plot.title = element_text(hjust = .5)) +
  facet_wrap(.~Primary.Type, scales = "free")

```
As we suspected earlier we can see that the most common location for gambling crimes is the street, sidewalk, and alley. All of which are outside therefore, during the winter less gamblind crimes will be committed. Another interesting note is the amoount of kidnapping crimes that happen in residence and apartment. This could be problems over custody of children or people breaking in to others homes. Now onto the geographical distribution of these crimes. 


#Crime Distribution
For our analysis of the Chicago Crime data set we will be focusing on the different distributions of crimes in chicago. Since there were over 30 different crimes, we chose 6 different crimes to analyze. These 6 crimes were chosen because of their distrubtion and the popularity of the crime. To best reperesed the distribution of all the crimes we chose to create heatmaps for all instances of each crime in the past 10 years. 



##Assault
```{r Assault, warning=FALSE}
singleCrime <- filter(hasLocation, Primary.Type =="ASSAULT")
ggmap(map, extent = "device") + geom_density2d(data = singleCrime, aes(x = Longitude, y = Latitude), size = 0.3) + 
  stat_density2d(data = singleCrime, 
                 aes(x = Longitude, y = Latitude, fill = ..level.., alpha = ..level..), size = 0.01, 
                 bins = 50, geom = "polygon") + scale_fill_gradient(low = "green", high = "red") + 
  scale_alpha(range = c(0, 0.3), guide = FALSE)
```

Assault shows a distribution of many smaller centers around Chicago. One interesting aspect of this heatmap is that downtown Chicago has a high rate of assaults which is not seen in many of the other crimes. Since assault can happen quickly and without planning this could lead to assaults appearting in more areas of Chicago.

##Gambling
```{r Gambling, warning=FALSE}

singleCrime <- filter(hasLocation, Primary.Type =="GAMBLING")
ggmap(map, extent = "device") + geom_density2d(data = singleCrime, aes(x = Longitude, y = Latitude), size = 0.3) + 
  stat_density2d(data = singleCrime, 
                 aes(x = Longitude, y = Latitude, fill = ..level.., alpha = ..level..), size = 0.01, 
                 bins = 50, geom = "polygon") + scale_fill_gradient(low = "green", high = "red") + 
  scale_alpha(range = c(0, 0.3), guide = FALSE)
```

Gambling shows a distribution very different from assault. The majority of gambling arrests are from one smaller area. This could be correlated to people running casinos out of their homes and simply moving it to a friends house after getting caught. Also people are willing to travel to gamble so other residents would likely travel to this area which allows underground casinos to cluster up in one area.

##Kidnapping
```{r Kidnapping, warning=FALSE}

singleCrime <- filter(hasLocation, Primary.Type =="KIDNAPPING")
ggmap(map, extent = "device") + geom_density2d(data = singleCrime, aes(x = Longitude, y = Latitude), size = 0.3) + 
  stat_density2d(data = singleCrime, 
                 aes(x = Longitude, y = Latitude, fill = ..level.., alpha = ..level..), size = 0.01, 
                 bins = 50, geom = "polygon") + scale_fill_gradient(low = "green", high = "red") + 
  scale_alpha(range = c(0, 0.3), guide = FALSE)
```

From this heatmap we can see that Kidnapping is very widespread throughout Chicago.  There is a large concentration to the West and South but overall the distribution seems large. This can mean that the kidnapping is more likely than other crimes to happen outside of just low income areas. However, there doesn't seem to be much of a presece in downtown Chicago, this is likely due to the fact that a kidnapping would be extremely difficult in a highly populated area.

##Robbery
```{r Robbery, warning=FALSE}

singleCrime <- filter(hasLocation, Primary.Type =="ROBBERY")
ggmap(map, extent = "device") + geom_density2d(data = singleCrime, aes(x = Longitude, y = Latitude), size = 0.3) + 
  stat_density2d(data = singleCrime, 
                 aes(x = Longitude, y = Latitude, fill = ..level.., alpha = ..level..), size = 0.01, 
                 bins = 50, geom = "polygon") + scale_fill_gradient(low = "green", high = "red") + 
  scale_alpha(range = c(0, 0.3), guide = FALSE)
```

Robbery is interesting becasue it seems to be clustered in smaller areas. This could be due to many factors such as areas of lower income or areas with more businesses with lower security. This shows an interesting comparison between robbery and assault, assault has larger centers of high concentrations while robbery has more smaller centers or high concentration. 

##Homicide
```{r Homicide, warning=FALSE}
singleCrime <- filter(hasLocation, Primary.Type =="HOMICIDE")
ggmap(map, extent = "device") + geom_density2d(data = singleCrime, aes(x = Longitude, y = Latitude), size = 0.3) + 
  stat_density2d(data = singleCrime, 
                 aes(x = Longitude, y = Latitude, fill = ..level.., alpha = ..level..), size = 0.01, 
                 bins = 50, geom = "polygon") + scale_fill_gradient(low = "green", high = "red") + 
  scale_alpha(range = c(0, 0.3), guide = FALSE)
```

Homicide shows a similar pattern to kidnapping where there are only 2 main clusters.  Since this is another more serious crime this could be due to people traveling to other areas to comit the crime or the difficulty of commiting the crime in very populated areas. Once again, it is also shown that the rate is low in downtown Chicago since the area is very populated and anyone who commits homicide there would certainly be caught.


##Narcotics
```{r Narcotics, warning=FALSE}
singleCrime <- filter(hasLocation, Primary.Type =="NARCOTICS")
ggmap(map, extent = "device") + geom_density2d(data = singleCrime, aes(x = Longitude, y = Latitude), size = 0.3) + 
  stat_density2d(data = singleCrime, 
                 aes(x = Longitude, y = Latitude, fill = ..level.., alpha = ..level..), size = 0.01, 
                 bins = 50, geom = "polygon") + scale_fill_gradient(low = "green", high = "red") + 
  scale_alpha(range = c(0, 0.3), guide = FALSE)
```

Narcotics shows a distribution very similar to gambling. This could be correlated to the majority of dealers living in one small area so arrests would all happen there. In this case we can see that West Chicago is the center of narcotics in Chicago.